Comment: Reasoning about Interference Between Units

### 2015-08-08 paper/introduction.tex
Working on final edit before submitting

\section{Background on randomization based statistical inference for questions
about causal effects from social network experiments}

In a randomized experiment with $n=4$ subjects connected via a fixed network,
subject $i=1$ might respond differently to different ways that treatment is
assigned to the whole network. When the treatment assignment vector provides
treatment to persons 2 and 3, $\bz=\{0,1,1,0\}$ she might act one way,
$y_{i=1,\bz=\{0,1,1,0\}}$ and when treatment is assigned to persons 3 and 4,
$\bz=\{0,0,1,1\}$ she might act another way, $y_{i=1,\bz=\{0,0,1,1\}}$ That
is, if the experiment had a causal effect on person $i$, then her outcome
would differ under different realizations of the experimental treatment
$y_{i,\bz} \ne y_{i,\bz'}$. The fundamental problem of causal inference
reminds us that we can never see both states of the world: we only observe the
outcome from person $i$ under one treatment assignment vector, either $\bz$ or
some $\bz'$ not both  \cite{holland:1986a}.

R.A. Fisher's approach to design-based statistical inference  \citep[Chap
2]{fisher:1935} as developed by Paul Rosenbaum  \citep{rosenbaum2010design}
begins with the premise of the fundamental problem of causal inference. Rather
than imagine that we could learn directly about the unobservable
counterfactual comparisons of outcomes under different treatment vectors,
Fisher would suggest that we focus on learning about how \emph{models} of
those counterfactual outcomes relate to what we observe. Although we do not
know how person $i$ would have acted under all possible experimental
scenarios, we can learn how much information we have to dispel certain claims
or hypotheses.  This conceptual move --- sidestepping the fundamental problem
of causal inference via learning about claims made by scientists --- drives
hypothesis testing in general.  BFP build on this insight by showing that
models of counterfactual effects can involve statements about how treatment
given to one node in a social network can come to have an effect on other
nodes. Their example model allows the effects of treatment to die off as the
network distance between nodes increases.\footnote{See the paper itself for
  details of the example model.} They further showed that the strength of
evidence against the specific hypotheses implied by a given model varied with
different features of the research design as well as the extent to which the
true causal process diverged from the model. Since their simulated experiment
involved only two treatments, the only observations available to evaluate the
model were comparisons of the assigned to treatment group and the assigned to
control group. Since their model could imply not only shifts in the mean of
the observed treatment versus control outcome distributions, but changes in
shape of those distributions, they used the Kolmogorov-Smirnov (KS) test
statistic so that their tests would be sensitive differences in the treatment
and control distributions implied by different hypotheses and not merely
sensitive to differences in one aspect of those distributions (such as the
differences in the mean).\footnote{If the empirical cumulative distribution
  function (ECDF) of the treated units is $F_1$ and the ECDF of the control
  units is $F_0$ then the KS test statistic is $\T(\yu,\bz) = \underset{i =
    1,\ldots,n}{\text{max}}\left[F_1(y_{i,\bzero}) -
    F_0(y_{i,\bzero})\right]$, where $F(x)=(1/n)\sum_{i=1}^n I(x_i \le x)$
  records the proportion of the distribution of $x$ at or below $x_i$
  \citep[\S 5.4]{MylesHollander1999a}.} So, in broad outline, the BFP approach
involves (1) the articulation of a model for how a treatment assignment vector
can change outcomes for all subjects in the experiment (holding the network
fixed) and (2) using function of comparing actually treated and control
observations to tell us whether such a model is implausible (codified as a low
$p$-value)  or whether we have too little information available from the data
and design about a the model (codified as a high $p$-value). This is classic
hypothesis testing.

>>>>

\section{Background on randomization based statistical inference for causal effects from social network experiments}

In a randomized experiment with $n=4$ subjects connected via a fixed network,
subject $i=1$ might respond differently to different ways that treatment is
assigned to the whole network. When the treatment assignment vector provides
treatment to persons 2 and 3, $\bz=\{0,1,1,0\}$ she might act one way,
$y_{i=1,\bz=\{0,1,1,0\}}$ and when treatment is assigned to persons 3 and 4,
$\bz=\{0,0,1,1\}$ she might act another way, $y_{i=1,\bz=\{0,0,1,1\}}$ That
is, if the experiment had a causal effect on person $i$, then her outcome
would differ under different realizations of the experimental treatment
$y_{i,\bz} \ne y_{i,\bz'}$. The fundamental problem of causal inference
reminds us that we can never see both states of the world: we only observe the
outcome from person $i$ under one treatment assignment vector, either $\bz$ or
some $\bz'$ not both  \citep{holland:1986a,brady2008cae}.

R.A. Fisher's approach to design-based statistical inference  \citep[Chap
2]{fisher:1935} as developed by Paul Rosenbaum  \citep{rosenbaum2010design}
begins with the premise of the fundamental problem of causal inference. Given
the unobservability of potential outcomes, the Fisher and Rosenbaum approach
suggests that we focus on learning about how \emph{models} of unobservable
counterfactual outcomes relate to what we can observe. Although we do not know
how person $i$ would have acted under all possible experimental scenarios, we
can learn how much information we have to dispel certain claims or hypotheses.
This conceptual move --- sidestepping the fundamental problem of causal
inference via learning about claims made by scientists --- drives hypothesis
testing in general.  BFP build on this insight by showing that models of
counterfactual effects can involve statements about how treatment given to one
node in a social network can influence other nodes. Their
example model allows the effects of treatment to die off as the network
distance between nodes increases.\footnote{See the paper itself for details of
  the example model.} They further show that the strength of evidence
against the specific hypotheses implied by a given model varies with different
features of the research design as well as the extent to which the true causal
process diverged from the model. Since their simulated experiment involved
two treatments, the only observations available to evaluate the model
were comparisons of the assigned to treatment group and the assigned to
control group. Since their model could imply not only shifts in the mean of
the observed treatment versus control outcome distributions, but changes in
shape of those distributions, they used the Kolmogorov-Smirnov (KS) test
statistic so that their tests would be sensitive to differences in the treatment
and control distributions implied by different hypotheses and not merely
sensitive to differences in one aspect of those distributions (such as the
differences in the mean).\footnote{If the empirical cumulative distribution
  function (ECDF) of the treated units is $F_1$ and the ECDF of the control
  units is $F_0$ then the KS test statistic is $\T(\yu,\bz) = \underset{i =
    1,\ldots,n}{\text{max}}\left[F_1(y_{i,\bzero}) -
    F_0(y_{i,\bzero})\right]$, where $F(x)=(1/n)\sum_{i=1}^n I(x_i \le x)$
  records the proportion of the distribution of $x$ at or below $x_i$
  \citep[\S 5.4]{MylesHollander1999a}.} So, in broad outline, the BFP approach
involves (1) the articulation of a model for how a treatment assignment vector
can change outcomes for all subjects in the experiment (holding the network
fixed) and (2) using function of comparing actually treated and control
observations to tell us whether such a model is implausible (codified as a low
$p$-value)  or whether we have too little information available from the data
and design about a the model (codified as a high $p$-value). This is classic
hypothesis testing.

### 2015-08-08 paper/paper.tex
Working on final edit before submitting

%%% Paper specific packages
\usepackage{setspace}
\usepackage[letterpaper,bottom=1in,top=1in,right=1in,left=1in,includemp=FALSE]{geometry}
\usepackage{microtype} %only useful with pdflatex
\usepackage{fancyvrb} % for verbatim text in footnotes

@@

% \ifanon
% \footnote{We are awesome!}
% \else
% \fi
%% \abovedisplayshortskip=0pt
%% \belowdisplayshortskip=0pt
%% \abovedisplayskip=0pt
%% \belowdisplayskip=0pt
%% 
\begin{document}
\VerbatimFootnotes

@@

\begin{abstract} 
\citet{bowers2013sutva} showed that one could use Fisher's randomization-based
hypothesis testing framework to assess counterfactual causal models of
treatment propagation and spillover across social networks.  This research
note improves the statistical infernce presented in \citet{bowers2013sutva}
(henceforth BFP) by substituting a test statistic based on a sum of squared
residuals and incorporating information about the fixed network for the simple
the Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
that they used. The point of this short article is to incrementally improve
the use of BFP's ``reasoning about interference'' approach. We do not offer
results about test statistics for multi-parameter causal models on social
networks here, but instead hope that this research note stimulates further,
and deeper, work on test statistics and sharp hypotheses.
\end{abstract}

@@

%\input{titlepage.tex}
%\setstretch{1.6} %1.6 or 1.37 which I think is "true double spacing"
%\onehalfspacing

>>>>

%%% Paper specific packages
\usepackage{setspace}
\usepackage[letterpaper,bottom=1in,top=1in,right=1in,left=1in,includemp=FALSE]{geometry}
%\usepackage{microtype} %only useful with pdflatex
%\usepackage{fancyvrb} % for verbatim text in footnotes

@@

% \ifanon
% \footnote{We are awesome!}
% \else
% \fi
%% \abovedisplayshortskip=0pt
%% \belowdisplayshortskip=0pt
%% \abovedisplayskip=0pt
%% \belowdisplayskip=0pt
%% 
\begin{document}
%\VerbatimFootnotes

@@

\begin{abstract} 
\citet{bowers2013sutva} showed that one could use Fisher's randomization-based
hypothesis testing framework to assess counterfactual causal models of
treatment propagation and spillover across social networks.  This research
note improves the statistical infernce presented in \citet{bowers2013sutva}
(henceforth BFP) by substituting a test statistic based on a sum of squared
residuals and incorporating information about the fixed network for the simple
Kolmogorov-Smirnov test statistic  \citep[\S 5.4]{MylesHollander1999a}
they used. This note incrementally improves
the application of BFP's ``reasoning about interference'' approach. We do not
offer general results about test statistics for multi-parameter causal models on social
networks here, but instead hope to stimulate further,
and deeper, work on test statistics and sharp hypothesis testing.
\end{abstract}

@@

%\input{titlepage.tex}
\setstretch{1.6} %1.6 or 1.37 which I think is "true double spacing"
%\onehalfspacing
